---
title: "team"
author: "Adam Brandt"
date: "9/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<font size="6"> **Reading the Data:**

```{r}
choco <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/data/choco.csv") 
str(choco) 
```

<font size="3">

5.	Brainstorm two analyses you would like to perform with your teammate. Describe in plain language what comparisons are you making. Create suitable numerical and graphical summaries for your analyses. Describe the results.

**Adam Brandt**

**Chocolate Rating by Location**

I wanted to see which location has the most high rated chocolate and which location has the lowest rated chocolate.
```{r}
GoodLoc <- choco[(choco$Rating >= 4) , ]
sort(GoodLoc$Location)
BadLoc <- choco[(choco$Rating <= 1) , ]
sort(BadLoc$Location)
barplot(table(choco$Location, choco$Rating))
```

The lowest rated chocolate comes from Belgium and the country with the most high rated chocolate is the U.S.A.

**Lisa McDaniel**

**USA Chocolate**

I got the summary statistics for the dataset and saw that there are 2312 records.  I filtered the original data to include only those records where the  *Location* was *U.S.A* and stored the 966 records in USAchoch.  I further filtered the data to include only those records where the  *Bean Origin* was also in the *U.S.A and stored the 28 records in USAbean.  I got the summary statistics for USABean and created plots showing the number of ratings per yeary and by cocoa percentage.

Ratings began in 2006 but have not been yearly.  Ratings peaked in 2012, 2013 and 2014 and have remained pretty steady over the past few years.

```{r echo=FALSE}
summary(choco)

USAchoc <- choco[(choco$Location == "U.S.A.") ,]
##USAchoc

USAbean <- USAchoc[(USAchoc$Bean_Origin == "U.S.A.") ,]
#USAbean
summary(USAbean)
#unique(USAbean$Date)
USAbeanCt1 <- table(USAbean$Date)
barplot(USAbeanCt1, main="USA: Ratings Over Time", names.arg= c("2006", "2009", "2010", "2011", "2012", "2013", "2014", "2016", "2019", "2020"), ylab="Number of Ratings", xlab="Year")
```

The most commonly rated chocolate has a cocoa percentage of 70%, followed by 72%.

```{r echo=FALSE}
##unique(USAbean$Cocoa_Percent)
USAbeanCt2 <- table(USAbean$Cocoa_Percent)
##summary(USAbean$Cocoa_Percent)
##USAbeanCt
barplot(USAbeanCt2, main="USA: Cocoa Percent", names.arg= c("55", "60", "65", "67", "70", "71", "72", "73", "75", "85"), ylab="Number of Ratings", xlab="Cocoa Percent")
          
```

Created box plots to show the ratings per cocoa percentage.  Chocolate with 67% and 75% have the most favorable rating of 3.75.

```{r echo=FALSE}
boxplot(USAbean$Rating~USAbean$Cocoa_Percent, main="USA: Cocoa Percent Rating", xlab="Cocao Percent", ylab = "Rating")
```


